---
title: "9. Support Vector Machines"
output: github_document
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = T, fig.align = "center")
```

- perfect separate: Maximal Margin Classifier, maximal margin hyperplane
- Support Vectors (points), Support Vector Classifier
- more flexible: Support Vector Machine
- kernel: linear kernel(alike Person correlation), polynomial kernel, radial kernel etc.
- bigger tuning parameter, many support vectors, wider margin, higher bias => low variance
- More than two classes, can use SVM
- almost (perfect) separate data set is uesful on SVM? 
- otherwise, Logistic Regression.
- but SVM exists perfect zero point.

### pre-requires
```{r warning = F, message = F}
library(e1071)      ## svm(), tune()
library(ROCR)       ## prediction(), performance()
data(Khan, package = "ISLR")
```

store unique data sets.
```{r eval = F}
appending = function(dataframe, string) {
  if(is.null(dim(dataframe)[1]))
    tag = rep(string, length(dataframe))
  else tag = rep(string, dim(dataframe)[1])
  dataframe = cbind(dataframe, tag)
  return(dataframe)
}

ytrain = appending(Khan$ytrain, "train")
ytest = appending(Khan$ytest, "test")

khan2 = cbind(rbind(Khan$xtrain, Khan$xtest), rbind(ytrain, ytest))
write.csv(khan2, "Khan.csv", row.names = F, quote = F)
```
- 처음에 너무 폭 넓게 만들고 조금씩 줄인 것. 딱히 함수 customizing은 할 필요 없었는데.

---

### 1. SVC: SV. Classifier (MMC, Maximal Margin Classifier)

```{r message = F}
## set.seed(1)
X = matrix(rnorm(20*2), ncol = 2)
y = c(rep(-1, 10), rep(1, 10))
X[y == 1,] = X[y == 1,] + 1
```

```{r}
plot(X, col = (3 - y))
```

```{r message = F}
dat = data.frame(X, y = as.factor(y))
svmfit = svm(y ~., dat, kernel = "linear", cost = 10, scale = F)
```
> default is radial kernel

- don't do standardization

```{r}
plot(svmfit, dat)
```

> plot.svm() no longer exist, maybe

```{r}
svmfit$index
summary(svmfit)
```

##### changing tuning parameter
```{r message = F}
## set.seed(1)
tune.out = tune(svm, y ~., data = dat, kernel = "linear", ranges = list(cost = c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
```
> tune(): 10-fold cross-validation

```{r}
summary(tune.out)
```
1 or 100 cost is best on my sampling data

```{r}
bestmod = tune.out$best.model
summary(bestmod)
```

##### test the made support vector classifier model
```{r message = F}
xtest = matrix(rnorm(20*2), ncol = 2)
ytest = sample(c(-1, 1), 20, rep = T)
xtest[ytest == 1,] = xtest[ytest == 1,] + 1
testdat = data.frame(xtest, y = as.factor(ytest))
```
- made confusion and conflict, because of colnames to upper alike sentense.

```{r}
ypred = predict(bestmod, testdat)
table(predict = ypred, truth = testdat$y)
```

##### test cost = 0.01
```{r}
svmfit = svm(y ~., dat, kernel = "linear", cost = 0.01, scale = F)
ypred = predict(svmfit, testdat)
table(predict = ypred, truth = testdat$y)
```

##### almost linearly separable, hyperplane
```{r}
X[y == 1,] = X[y == 1,] + .5
plot(X, col = (y + 5)/2, pch = 19)
```

```{r}
dat = data.frame(X, y = as.factor(y))
svmfit = svm(y ~., dat, kernel = "linear", cost = 1e5)
summary(svmfit)
```
- set gamma = .5, but don't changed
- still use many support vectors, like 8
- set and use very randomly seed?

```{r}
svmfit = svm(y ~., dat, kernel = "linear", cost = 1)
summary(svmfit)
plot(svmfit, dat)
```

---

### 2. SVM: Support Vector Machine

##### almost non-linear
```{r message = F}
## set.seed(1)
X = matrix(rnorm(200*2), ncol = 2)
X[1:100,] = X[1:100,] + 2
y = c(rep(1, 150), rep(2, 50))
dat = data.frame(X, y = as.factor(y))
```

```{r}
plot(X, col = y)
```

- col = c("blue", "green"), pretty good

```{r message = F}
train = sample(200, 100)
svmfit = svm(y ~., dat[train,], gamma = 1, cost = 1)
```
> default is radial kernel

> naturally, exist difference with use custom gamma and default gamma

```{r}
plot(svmfit, dat[train,])
summary(svmfit)
```

```{r}
svmfit = svm(y ~., dat[train,], gamma = 1, cost = 1e5)
plot(svmfit, dat[train,], col = c("azure", "lightblue"))
```

##### change tuning parameter
```{r}
## set.seed(1)
tune.out = tune(svm, y ~., data = dat[train,], ranges = list(cost = c(.1, 1, 10, 100, 10^3), gamma = c(.5, 1, 2, 3, 4)))
summary(tune.out)
```

```{r}
table(true = dat[-train, 'y'], pred = predict(tune.out$best.model, newdata = dat[-train,]))
```
- "true", "pred" are labeling on table

misclassifing ratio is 28%

---

### ROC Curves

```{r message = F}
rocplot = function(pred, truth, ...) {
  predob = prediction(pred, truth)
  perf = performance(predob, "tpr", "fpr")
  plot(perf, ...)
}
```

##### train data
```{r message = F}
svmfit.opt = svm(y ~., dat[train,], gamma = 2, cost = 1, decision.values = T)
train.fitted.opt = attributes(predict(svmfit.opt, dat[train,], decision.values = T))$decision.values
```
- best is gamma = 2, cost = 1 by book

```{r message = F}
svmfit.flex = svm(y ~., dat[train,], gamma = 50, cost = 1, decision.values = T)
train.fitted.flex = attributes(predict(svmfit.flex, dat[train,], decision.values = T))$decision.values
```

##### test data
```{r message = F}
test.fitted.opt = attributes(predict(svmfit.opt, dat[-train,], decision.values = T))$decision.values
test.fitted.flex = attributes(predict(svmfit.flex, dat[-train,], decision.values = T))$decision.values
```

```{r}
par(mfrow = c(1,2))
rocplot(train.fitted.opt, dat[train, 'y'], main = "Training Data")
rocplot(train.fitted.flex, dat[train, 'y'], add = T, col = "red")
rocplot(test.fitted.opt, dat[-train, 'y'], main = "Test Data")
rocplot(test.fitted.flex, dat[-train, 'y'], add = T, col = "red")
```

---

### SVM with Multiple classes

```{r message = F}
## set.seed(1)
X = rbind(X, matrix(rnorm(50*2), ncol = 2))
y = c(y, rep(0, 50))
X[y == 0, 2] = X[y == 0, 2] + 2
dat = data.frame(X, y = as.factor(y))
```

```{r}
plot(X, col = (y + 1))
```

```{r}
svmfit = svm(y ~., dat, cost = 10, gamma = 1)
plot(svmfit, dat, col = c("lightblue", "azure", "lightgreen"))
```
- changed colors, first colors so messy

---

### Application to Gene Expression Data
- functions are very normal functioning

```{r}
names(Khan)
dim(Khan$xtrain); dim(Khan$xtest)
length(Khan$ytrain); length(Khan$ytest)
table(Khan$ytrain); table(Khan$ytest)
```

```{r message = F}
dat = data.frame(Khan$xtrain, y = as.factor(Khan$ytrain))
```

```{r}
out = svm(y ~., dat, kernel = "linear", cost = 10)
summary(out)
table(out$fitted, dat$y)
```

```{r}
dat.te = data.frame(Khan$xtest, y = as.factor(Khan$ytest))
pred.te = predict(out, newdata = dat.te)
table(pred.te, dat.te$y)
```
