---
title: "3. Linear Regression"
output: github_document
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = T)
```

### pre-requires
```{r include = F}
## library(MASS)
library(car)    ## vif()
data(Boston, package = "MASS")
data(Carseats, package = "ISLR")
```

store unique data sets.
```{r eval = F}
write.csv(Boston, "Boston.csv", row.names = F, quote = F)
write.csv(Carseats, "Carseats.csv", row.names = T, quote = F)
```
---
### 1. Simple Linear Regression
data checking 
```{r}
## fix(Boston)
names(Boston)
```
> fix( ): whole check for data set

```{r}
lm.fit = lm(medv ~ lstat, data = Boston)
## attach(Boston);	lm.fit = lm(medv ~ lstat)
summary(lm.fit);	summary(lm.fit)$r.sq;	summary(lm.fit)$sigma
## names(lm.fit);	coef(lm.fit);
```
confint(lm.fit): coefficient 2.5, 97.5 percentile value

```{r}
predict(lm.fit, data.frame(lstat = c(5,10,15)), interval = "confidence")
predict(lm.fit, data.frame(lstat = c(5,10,15)), interval = "prediction")
```

```{r}
plot(Boston$lstat, Boston$medv)
abline(lm.fit)
## abline(lm.fit, lwd = 3);         abline(lm.fit, lwd = 3, col = "red")
## plot(lstat, medv, col = "red");  plot(lstat, medv, pch = 20)
## plot(lstat, medv, pch = "+");	plot(1:20, 1:20, pch = 1:20)
```

```{r}
par(mfrow = c(2,2))
plot(lm.fit)
```

```{r}
par(mfrow = c(1,3))
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
## external studentization residuals
plot(hatvalues(lm.fit))
```

```{r}
which.max(hatvalues(lm.fit))
```
> which.max( ): return maximum data


---
### 2. Multiple Linear Regression
```{r}
lm.fit = lm(medv ~ lstat + age, data = Boston)
summary(lm.fit)
## lm.fit = lm(medv ~, data = Boston)
## summary(lm.fit)
```

```{r}
vif(lm.fit)
```

```{r}
lm.fit1 = lm(medv ~. - age, data = Boston)
## lm.fit1 = update(lm.fit, ~.-age)
summary(lm.fit1)
```
> update( ): literally update (fit) function

---
### 3. Interaction Term
```{r}
summary(lm(medv ~ lstat*age, data = Boston))
```
- '*' means adding all (two) variables and interaction term
- ':' means adding first variable and interaction term

---
### 4. Non-linear Transormations of the Predictors
```{r}
lm.fit2 = lm(medv ~ lstat + I(lstat^2), data = Boston)
summary(lm.fit2)
```
represent two-square term through '^', 'I( )'

```{r}
par(mfrow = c(2,2))
plot(lm.fit2)
```

##### compare with normal and quadratic
```{r}
lm.fit = lm(medv ~ lstat, data = Boston)
anova(lm.fit, lm.fit2)
```

##### 5-degree poly
```{r}
lm.fit5 = lm(medv ~ poly(lstat, 5), data = Boston)
summary(lm.fit5)
```

##### log
```{r}
summary(lm(medv ~ log(rm), data = Boston))
```

---
### 5. Qualitative Predictors
data checking 
```{r}
## fix(Carseats)
names(Carseats)
```

```{r}
lm.fit = lm(Sales ~. + Income:Advertising + Price:Age, data = Carseats)
summary(lm.fit)
```

```{r eval = F}
contrasts(Carseats$ShelveLoc)
```
> contrasts( ): making dummy variables

---
boldly skip for custom (combine) function
